---
title: "Plots for publication"
output: html_document
  html_document:
    theme: paper
    highlight: zenburn
    number_sections: true
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: true
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This R markdown file creates the plots for our paper on SpecEnr.

flowGraph takes flowType Phenotypes cell populations as input and outputs SpecEnr values for each cell populations nodes (immunophenotypes), within a flow cytometry sample, based on their expected proportion. In this way, SpecEnr accounts for relations between cell populations to produce cell population quantification whose changes are not induced by neighbouring cell populations but represent real differential abundance behaviour.

## Citation

The theory, proof, and algorithm behind the SpecEnr statistic used in the flowGraph package can be found in the following [paper](https://www.biorxiv.org/content/10.1101/837765v3.abstract). Please consider citing if you found it helpful.

bibtex:
```
@article{yue2019identifying,
  title={Identifying differential cell populations in flow cytometry data accounting for marker frequency},
  author={Yue, Alice and Chauve, Cedric and Libbrecht, Maxwell and Brinkman, Ryan},
  journal={BioRxiv},
  pages={837765},
  year={2019},
  publisher={Cold Spring Harbor Laboratory}
}
```

## Data, working directory, & packages

Extract the .Rdata files coupled packaged with this script into the working directory of this R markdown file.

```{r}
# load packages
# install flowGraph using devtools if not already installed
# library(devtools)
# devtools::install_github("aya49/flowGraph")
library(flowGraph)
library(lattice)

# set working directory if not done already
root <- "/mnt/f/Brinkman group/current/Alice/flowtype_metric/pub" 
setwd(root)
```

# Statistic robustness plots

Negative control ctrl data set: T-test -ln(p-value) qq plot.

```{r}
load("data/ctrl.Rdata")

# raw cell count
fg_plot_qq(fg, index=1, adjust_custom="none", logged=TRUE)

# proportion
fg_plot_qq(fg, index=3, adjust_custom="none", logged=TRUE)

# SpecEnr
fg_plot_qq(fg, index=2, adjust_custom="none", logged=TRUE)
```

Negative control ctrl data set: proportion of significant T-test p-value <.05.

```{r}
pp <- fg_get_summary(fg, index=1, adjust_custom="none")

# total number of cell populations
cpopl <- length(pp$values)

# proportion of significant cell populations
sig_count <- sum(pp$values<.05)
pp <- fg_get_summary(fg, index=3, adjust_custom="none")
sig_prop <- sum(pp$values<.05)
pp <- fg_get_summary(fg, index=2, adjust_custom="none")
sig_se <- sum(pp$values<.05)

(cpopl)
(sig_count)
(sig_prop)
(sig_se)
```

Positive control pos1 data set: F-measure related statistics measuring consistency of p-values between replicates of pos1

```{r}
load("data/pos11.Rdata")

# calculate p-values
p_adjust <- function(X, inds) {
    p <- apply(as.matrix(X), 2, function(x)
        tryCatch(t.test(x[inds[[1]]],x[inds[[2]]])$p.value, error=function(e)1))
    p[is.na(p)] <- 1
    return(p.adjust(p, method="bonferroni"))
}

inds1 <- list(c(1:5), c(11:15))
inds2 <- list(c(6:10), c(16:20))
p_se1 <- p_adjust(fg@feat$node$SpecEnr, inds=inds1)
p_se2 <- p_adjust(fg@feat$node$SpecEnr, inds=inds2)
p_count1 <- p_adjust(fg@feat$node$count, inds=inds1)
p_count2 <- p_adjust(fg@feat$node$count, inds=inds2)
p_prop1 <- p_adjust(fg@feat$node$prop, inds=inds1)
p_prop2 <- p_adjust(fg@feat$node$prop, inds=inds2)

# calculate statistics
p_stats <- function(p1, p2, pthres=.05) {
    p1w <- which(p1<pthres)
    p2w <- which(p2<pthres)
    p12w <- intersect(p1w, p2w)
    
    prec <- length(p12w) / length(p1w)
    reca <- length(p12w) / length(p2w)
    
    p_stats <- list(precision=prec, recall=reca,
                f1=2*prec*reca/(prec+reca))
    return(unlist(p_stats))
}

sbs <- data.frame(score=c(p_stats(p_se1, p_se2), 
                          p_stats(p_count1, p_count2), 
                          p_stats(p_prop1, p_prop2)),
                  feature=rep(c("SpecEnr","count","prop"),each=3),
                  metric=rep(c("precision","recall","f1"),times=3))

# plot barplot
pl = lattice::barchart(score~feature, sbs, groups=metric, las=1, ylim=c(0,1),
                       auto.key = list(space = "top"),
                       scales=list(x=list(rot=45)), 
                       cex.lab=2, cex.main=2, cex.sub=2, cex.axis=1.5)
print(pl)
```

# Cell hierarchy plots (SpecEnr vs prop q-values)

All coloured nodes have a significant q-value.

Only a limited number of labels are displayed to reduce clutter.

```{r}
ch_plot <- function(fg, index, adjust_custom="byLayer", pthres=.05,
                    show_bgedges=TRUE, truepos=c()) {
    # sm <- unlist(fg@summary_desc$node[index,])
    # pp1 <- fg_get_summary(
    #     fg, index=index,adjust_custom=adjust_custom,
    #     filter_btwn_tpthres=.05, filter_btwn_es=.5)
    # pp1sig <- pp1$values<pthres
    # sort(pp1$values[pp1sig])
    # sum(pp1sig)
    
    gr <- fg_plot(
        fg, p_thres=pthres, index=index,
        show_bgedges=show_bgedges,
        interactive=FALSE, 
        adjust_custom=adjust_custom,
        filter_btwn_tpthres=.05, filter_btwn_es=.5,
        node_labels="NONE")
    gr$v$label <- gr$v$phenotype
    
    
    gr$v$label_ind <- FALSE
    gr$v$label_ind[gr$v$phenotype%in%truepos] <- TRUE
    gr_ <- gr
    
    gp <- plot_gr(gr_, label_coloured=FALSE)
    return(gp)
}
```

## Data set: pos1 (artificially increased/decreased); SpecEnr vs prop

```{r}
truepos <- c("A+","A-")

load("data/pos11.Rdata")
plot(ch_plot(fg, index=2, truepos=truepos))
plot(ch_plot(fg, index=1, truepos=truepos))
load("data/pos31.Rdata")
plot(ch_plot(fg, index=2, truepos=truepos))
plot(ch_plot(fg, index=1, truepos=truepos))
```

## Data set: pos2 (artificially increased/decreased); SpecEnr vs prop

```{r}
truepos <- c("A+B+C+","A+B+","A+C+","B+C+","A+","B+","C+")

load("data/pos15.Rdata")
plot(ch_plot(fg, index=2, truepos=truepos))
plot(ch_plot(fg, index=1, truepos=truepos))
load("data/pos35.Rdata")
plot(ch_plot(fg, index=2, truepos=truepos))
plot(ch_plot(fg, index=1, truepos=truepos))
```

## Data set: pos3 (artificially increased/decreased); SpecEnr vs prop

```{r}
truepos <- c("A+","B+","D+","A+B+","A+B+D+")

load("data/pos26.Rdata")
plot(ch_plot(fg, index=2, truepos=truepos))
plot(ch_plot(fg, index=1, truepos=truepos))
load("data/pos36.Rdata")
plot(ch_plot(fg, index=2, truepos=truepos))
plot(ch_plot(fg, index=1, truepos=truepos))
```

## Data set: flowcap compares AML vs healthy samples; SpecEnr vs prop

```{r}
truepos <- c("FS-","SS-","HLA+","CD38+",
             "FS-CD34+","SS-CD34+",
             "HLA+CD117-CD45+CD34+",
             "FS-SS+CD45-","FS-SS+CD117+CD45-")

load("data/flowcap.Rdata")
plot(ch_plot(fg, index=3, truepos=truepos, show_bgedges=FALSE))
plot(ch_plot(fg, index=1, truepos=truepos, show_bgedges=FALSE))
```

## Data set: pregnancy compares late-term pregnancy vs post-partum blood samples; SpecEnr vs prop

```{r}
truepos <- c("CD3+","CD45RA+","CD16+",
             "CD3+CD45RA+CD56-Tbet-",
             "CD3+CD45RA+CD56-CD8-Tbet-TCRgd-",
             "CD3+CD45RA+CD56-CD7+CD8-Tbet-",
             "CD3+CD45+CD45RA+CD56-CD8-Tbet-",
             "CD3+CD45RA+CD45-CD8-Tbet-",
             "CD7+CD8+","CD7+CD8+Tbet+",
             "CD16+CD56-CD7-TCRgd-")

load("data/pregnancy.Rdata")
plot(ch_plot(fg, index=6, truepos=truepos, show_bgedges=FALSE))
plot(ch_plot(fg, index=3, truepos=truepos, show_bgedges=FALSE))
```
